Correcting for bias in high-throughput sequencing data

### 2011-09-23 seqbias-paper.tex
changes

Because we cannot observe directly the underlying RNA-abundance, our evaluation
strategy relies on testing two assumptions we make of an ideal, unbiased experiment.
\begin{enumerate}
\item Positional nucleotide frequencies (as in Figure \ref{fig:freqs}), measured
from reads within exons, should not differ greatly from frequencies measured by
sampling uniformly within the same exons.
\item Read counts across a single exon should follow, approximately, a Poisson
process.
\end{enumerate}

@@

>>>>

Because we cannot observe directly the underlying RNA-abundance, our evaluation
strategy relies on testing three assumptions we make of an ideal, unbiased experiment.
\begin{enumerate}
\item Positional nucleotide frequencies (as in Figure \ref{fig:freqs}), measured
from reads within exons, should not differ greatly from frequencies measured by
sampling uniformly within the same exons.
\item Read counts across a single exon should follow, approximately, a Poisson
process.
\item Adjusting for bias should increase the agreement between RNA-Seq and
qRT-PCR.
\end{enumerate}

@@

\subsection{qRT-PCR Correlation}

We used sequencing data previously published by \citet{Au2010} to evaluate the
effect bias correction has on correlation to measurements made by TaqMan RT-PCR,
made available by the the Microarray Quality Control project \cite{Shi2006}. The
RNA-Seq data shows a pattern of bias similar to that seen in the other samples
sequenced on an Illumina platform.
%% TODO: refer to supplement for figure of bias

To evaluate the efficacy of each of the bias correction methods considered, we
counted reads overlapping each gene, defining the gene by the union of every
transcript in release 62 of the Ensembl gene annotations, excluding UTRs, and
100 additional bases from both ends of the gene. Counts were then normalized
(divided) by the length of these genes. We then removed any genes with a read
count less than 10, or that did not correspond to a unique TaqMan probe, leaving
592 genes.

%% TODO: something goes here
After adjusting read counts according to the predicted sequence bias, we
computed the Pearson correlation between log read counts and log TaqMan
expression values, which are averaged across three replicates.

\begin{figure}
\centerline{\includegraphics[width=0.4\textwidth]{pcrcor.eps}}
\caption{
    Pearson correlation is computed betweer TaqMan values and (bias-corrected)
    read counts, on a logarithmic scale.
    %%TODO
    Error bars show a 95\% confidence interval, estimated using the Fisher
    transformation, as computed by the R function \texttt{cor.test}.
}
\label{fig:pcrcor}
\end{figure}

%\begin{table}
%\processtable{
    %Pearson correlation between log-adjusted read counts and log-adjusted TaqMan
    %values.\label{tab:pcr}}{
%\begin{tabular}{ll} \toprule
%Method & Correlation \\ \midrule
%Unadjusted &  0.6665 \\
%7mer &  0.6677 \\
%GLM & 0.6854 \\
%MART & 0.6921 \\
%BN & 0.7016
%\end{tabular}
%}
%{
%}
%\end{table}
